Differences in Homicide Case Information Indicates Why Justice is Not Served

### 2024-12-03 paper/paper.qmd
Updated paper qmd file to fix commas and update wording

The raw dataset we retrieved from The Washington Post using a script that downloads the CSV file from their GitHub repository contains 52179 observations with one observation being a homicide case. However, since we narrowed our scope to focus on the two of the most populated US cities, Los Angeles and Chicago, the number of observations in our cleaned dataset ended up being 6307. Our data look as follows:

@@

@tbl-dataset-preview and @tbl-dataset-statistics indicates our variables of interest, which are the following: victim_race, victim_age, victim_sex, city, disposition, year, month, and arrest_was_not_made. victim_race represents the race of the homicide victim, which can be "White" (376 observations), "Hispanic" (1763 observations), "Black" (4063 observations), "Asian" (43 observations), and "Other" (62 observations). victim_age signifies the age of the homicide victim at the time of their death and is defined as a whole number. @tbl-dataset-statistics reveals that the mean victim_sex is the sex of the homicide victim where they are either identified as a "female" or "male". The city variable defines the city the victim is reported to have been found in. The disposition variable is the specific status of a homicide case where a case can fall in either of the following three status: "Closed by arrest", "Open/No arrest", and "Closed without arrest". The year variable represents a year from 2010 to 2017 that indicates the year the homicide took place. Following this, the month variable represents the month a homicide took place. arrest_was_not_made is a variable that was constructed based on the disposition variable with "Closed by arrest" being converted to a 0 and "Open/No arrest" and "Closed without arrest" being converted to a 1. arrest_was_not_made indicates the status of a homicide case as either being unsolved, which is denoted by a 1, and solved, which is denoted by a 0.

@@

@fig-city-bar and @tbl-city-table shows that 77% of Chicago's homicide cases are unsolved (with 3164 cases) while for Los Angeles, 49% (1087 cases) of their homicide cases are unsolved from 2010 to 2017. Only 23% (947 cases) of homicide cases are solved with an arrest made for Chicago while for Los Angeles, 51% (1109 cases) of homicides are solved with an arrest made. Chicago has a higher percentage and higher number of unsolved homicide cases compared to Los Angeles.

@@

@fig-disposition-bar and @tbl-disposition-table shows that the number of homicide cases that are closed by arrest at 1109 cases is close to the number of homicide cases that are open at 1087 cases from 2010 to 2017. On the hand, there is a larger gap for Chicago between the number of homicides that are closed by arrest at 947 cases with the ones that are closed without arrest and open/no arrest at 216 and 2948 cases, respectively.

@@

@fig-victim-sex-bar and @tbl-victim-sex-table shows that most victims of homicide are male with 1721 solved cases and 3903 unsolved cases in comparsion to female victims with 335 solved cases and 348 unsolved cases. The proportion of cases with arrest and not is almost equal for homicides with female victims at 49% and 51%, respectively. On the other hand, the proportion of cases with arrest (31%) and not (69%) has a 38% difference for homicides with female victims.

@@

From @fig-victim-race-bar and @tbl-victim-race-table, they show that there were more homicide case solved when the victims identified as "White" (191 cases), "Asian" (22 cases), and "Other" (41 cases). However, a disproportionate number of homicide cases have victims who are Black or Hispanic. For both Black and Hispanic victims, they have more homicide cases that go unsolved with 2960 and 1064 cases compared to homicide cases that are solved with 1103 and 699 cases, respectively.

@@

**Demographic Questions**
The purpose of these questions is to gauge the respondents' educational background. Questions that would be asked includes information about their email, age, gender, ranking on the team, and educational attainment.

**Job-related Questions**
These questions will mostly consist of a mix of single and multiple-choice questions and open-ended questions where we will be asking about their feelings and experience handling homicide cases.

@@

Contact Information: 
- Name: Emily Su
- Email: em.su\@mail.utoronto.ca

**Demographic Questions**
1. What is your email? (Open-ended question)
2. How old are you? (Open-ended question but can only input numbers greater than or equal to 0)
3. Which gender do you identify with? (Single-choice question with option to enter gender if selected "Other") Female, Male, Other:___ (Have the option to input gender they identify with)
4. What is your ranking on your team? ____ (Open-ended question)
5. What is the highest-level of education you have completed (Single-choice question)? No formal education/kindergarten only, Elementary, Middle School, High School, General Equivalency Diploma (GED) or equivalent, Some college, Associate's Degree, Bachelor's Degree, Master's Degree, Professional degree, Doctoral degree

>>>>

The raw dataset we retrieved from The Washington Post using a script that downloads the CSV file from their GitHub repository contains 52,179 observations with one observation being a homicide case. However, since we narrowed our scope to focus on the two of the most populated US cities, Los Angeles and Chicago, the number of observations in our cleaned dataset ended up being 6,307. Our data look as follows:

@@

@tbl-dataset-preview and @tbl-dataset-statistics indicates our variables of interest, which are the following: victim_race, victim_age, victim_sex, city, disposition, year, month, and arrest_was_not_made. victim_race represents the race of the homicide victim, which can be "White" (376 observations), "Hispanic" (1,763 observations), "Black" (4,063 observations), "Asian" (43 observations), and "Other" (62 observations). victim_age signifies the age of the homicide victim at the time of their death and is defined as a whole number. @tbl-dataset-statistics shows that the mean victim_sex is the sex of the homicide victim where they are either identified as a "female" or "male". The city variable defines the city the victim is reported to have been found in. The disposition variable is the specific status of a homicide case where a case can fall in either of the following three status: "Closed by arrest", "Open/No arrest", and "Closed without arrest". The year variable represents a year from 2010 to 2017 that indicates the year the homicide took place. Following this, the month variable represents the month a homicide took place. arrest_was_not_made is a variable that was constructed based on the disposition variable with "Closed by arrest" being converted to a 0 and "Open/No arrest" and "Closed without arrest" being converted to a 1. arrest_was_not_made indicates the status of a homicide case as either being unsolved, which is denoted by a 1, and solved, which is denoted by a 0.

@@

@fig-city-bar and @tbl-city-table shows that 77% of Chicago's homicide cases are unsolved (with 3,164 cases) while for Los Angeles, 49% (1,087 cases) of their homicide cases are unsolved from 2010 to 2017. Only 23% (947 cases) of homicide cases are solved with an arrest made for Chicago while for Los Angeles, 51% (1,109 cases) of homicides are solved with an arrest made. Chicago has a higher percentage and higher number of unsolved homicide cases compared to Los Angeles.

@@

@fig-disposition-bar and @tbl-disposition-table shows that the number of homicide cases that are closed by arrest at 1,109 cases is close to the number of homicide cases that are open at 1,087 cases from 2010 to 2017. On the hand, there is a larger gap for Chicago between the number of homicides that are closed by arrest at 947 cases with the ones that are closed without arrest and open/no arrest at 216 and 2,948 cases, respectively.

@@

@fig-victim-sex-bar and @tbl-victim-sex-table shows that most victims of homicide are male with 1,721 solved cases and 3,903 unsolved cases in comparsion to female victims with 335 solved cases and 348 unsolved cases. The proportion of cases with arrest and not is almost equal for homicides with female victims at 49% and 51%, respectively. On the other hand, the proportion of cases with arrest (31%) and not (69%) has a 38% difference for homicides with female victims.

@@

From @fig-victim-race-bar and @tbl-victim-race-table, they show that there were more homicide case solved when the victims identified as "White" (191 cases), "Asian" (22 cases), and "Other" (41 cases). However, a disproportionate number of homicide cases have victims who are Black or Hispanic. For both Black and Hispanic victims, they have more homicide cases that go unsolved with 2,960 and 1,064 cases compared to homicide cases that are solved with 1,103 and 699 cases, respectively.

@@

**Demographic Questions**:

The purpose of these questions is to gauge the respondents' educational background. Questions that would be asked includes information about their email, age, gender, ranking on the team, and educational attainment.

**Job-related Questions**:

These questions will mostly consist of a mix of single and multiple-choice questions and open-ended questions where we will be asking about their feelings and experience handling homicide cases.

@@

Contact Information:

- Name: Emily Su

- Email: em.su\@mail.utoronto.ca

**Demographic Questions**
1. What is your email? (Open-ended question)

2. How old are you? (Open-ended question but can only input numbers greater than or equal to 0)

3. Which gender do you identify with? (Single-choice question with option to enter gender if selected "Other") Female, Male, Other:___ (Have the option to input gender they identify with)

4. What is your ranking on your team? ____ (Open-ended question)

5. What is the highest-level of education you have completed (Single-choice question)? No formal education/kindergarten only, Elementary, Middle School, High School, General Equivalency Diploma (GED) or equivalent, Some college, Associate's Degree, Bachelor's Degree, Master's Degree, Professional degree, Doctoral degree
